---
title: "Thing 2"
format: gfm
---

## Installation

You can install the development version of thing2 from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("nxskok/thing2")
```

## The data

The data on which this mini-analysis was based came from here:

```{r}
library(Devore7)
data(ex15.25)
ex15.25
```

In Devore's textbook it says, about these data:

The accompanying data on cortisol level was reported in the
article “Cortisol, Cortisone, and 11-Deoxycortisol Levels in
Human Umbilical and Maternal Plasma in Relation to the
Onset of Labor” (J. Obstetric Gynaecology British Com-
monwealth, 1974: 737–745). Experimental subjects were
pregnant women whose babies were delivered between 38
and 42 weeks gestation. Group 1 individuals elected to
deliver by Caesarean section before labor onset, group 2
delivered by emergency Caesarean during induced labor,
and group 3 individuals experienced spontaneous labor.

This indicates that each pregnant woman contributed one birth to the dataset, and we
have independent observations. This is not the impression conveyed by this
display of the dataset, which was evidently laid out this way to display in the 
textbook rather than being ready for analysis.

The purpose of this package is to illustrate the use of writing a package for a simple analysis.

The function `read_raw` reorganizes these data to be suitable for some kind of analysis
of variance. In particular:

- the dataframe is pivoted longer
- the missing values are removed (these are "structural" missing values, not actual missing data)
- the group numbers are replaced by names that indicate which group is which:

```{r}
library(thing2)
births <- read_raw()
births
```

The levels of `type` indicate whether the labour was spontaneous (`spont`), whether a Caesarean section was 
planned before labour started (`c_pre`), or whether an emergency Caesarian section was used to deliver the baby during induced labour (`c_emerg`). Interest in this study was in whether the mother's average cortisol level differed between the three types of birth.

## Making a graph

The function `draw_boxplot`, as the name suggests, draws a boxplot to compare the distributions of cortisol levels between the three groups of mothers:

```{r}
draw_boxplot(births)
```

The median cortisol levels clearly differ among the three groups. In addition, the shapes of the distributions differ
substantially, with the spontaneous-labour births having much the greatest variability. There are also outliers
in the `c_pre` group, though one might argue that these appear as outliers
 because of the small variability in the other observations.

## Analysis

The graph reveals that we should not consider
 these two analyses:
 
- a standard ANOVA because the groups differ substantially in spread
- despite the outliers, a Kruskal-Wallis test is not appropriate because the distributions differ in shape (the assumption is sometimes expressed as "the distributions differ only in centre").

The best parametric analysis is a Welch ANOVA (done by `oneway.test`), and the appropriate followup is the Games-Howell test (done by `gamesHowellTest` in package `PMCMRplus`), as done by function `run_welch`:

```{r}
run_welch(births)
```

The first output is the overall Welch test, which is significant, so the mean cortisol levels are not all equal. The second one 
shows which groups differ significantly: just the two Caesarean groups. You might find this surprising, because the median of `spont` is much the largest, but the `spont`  values are much more spread out, so that it is not possible to prove that the mean of this group is different from the means of either of the other two groups. The two Caesarean groups *are* significantly different in mean cortisol, however, because these groups have much smaller spread than the `spont` group.

If the outliers are a concern to you, the appropriate non-parametric test is Mood's median test, which makes no assumptions about variances. I wrote a package `smmr` (not on CRAN) to run this test:

```{r}
# install_github("nxskok/smmr")
library(smmr)
median_test(births, cortisol, type)
```

This has a similar significance to the Welch test (with a larger P-value), and indeed a possible followup (all possible Mood's median tests with a Bonferroni correction) has a similar result to the Games-Howell test:

```{r}
pairwise_median_test(births, cortisol, type)
```

that is to say, once again only the two Caesarian groups differ significantly in (now) median.

